---
title: "Population Densities - Comparing Across (Selected) Countries"
output:
  html_document:
    df_print: paged
    css: "style.css"
  html_notebook: default
keep_md: yes
params:
  ccode: "Country_Code"
  sabbr: "State_Abbr"
  msa_name: "MSA_Name"
  msa_code: "MSA_Code"
  sa_code: "SA_Code"
  ppn_est: "Popn_Est"
  area: "Area_km2"
  density: "Density"
  ppn_pct: "Perc_Popn"
  cum_pct: "Cum_Popn_Perc"
  method: "method"
  table_css: "table"
---

## Notebook Summary 
This notebook looks at population density metrics for a set of selected countries.
The basic flow is reading in the data, some data manipulation, then reviewing
various plots

Current countries in the data:
* Australia
* United States of America

---

### 1. Initialise and Read in the Data
Data is sourced from the relevant national statistical agencies and gathered
using scripts in the `data-raw` directory with results stored as `csv` files in
the `inst/extdata` directory and as `rda` files in the `data` directory.
<br>  

```{r setup}
library(geo.popn.density)
InitCSS()

all_data <- GetAllData()
```
<br>  

### 2. Data Review
#### 2.1 Data Coverage
##### Total Population, Area and Avg. Density by Country
```{r datacoverage}
ShowTable(TableCountryTotals(all_data), params$table_css)
```
<br>  

##### Population Summary Statistics by Statistical Area
```{r sapopnstats}
ShowTable(TableSummStats(all_data, stat_col = params$ppn_est), params$table_css)
```
<br>  

##### And by Metropolitan Area
``` {r msapopnstats}
all_data_exc_rural <- ExcRural(all_data, msa_name = params$msa_name, msa_code = params$msa_code)
ShowTable(TableSummStats(all_data_exc_rural, group_by = params$msa_name, stat_col = params$ppn_est), params$table_css)
```
<br>  

##### Area Summary Statistics by Statistical Area
``` {r saareastats}
ShowTable(TableSummStats(all_data, stat_col = params$area, dflt_dp = 2), params$table_css)
```
<br>  

##### And by Metropolitan Area
``` {r msaareastats}
ShowTable(TableSummStats(all_data_exc_rural, group_by = params$msa_name, stat_col = params$area, dflt_dp = 2), params$table_css)
```
<br>  

#### Discussion 2022-06-21
The "Statistical Area Level 2" is the lowest level statistical area data
generally available for Australian population and is driving a much lower
fidelity than that available for the US. This leads to inevitable issues
of just how comparable these statistic actually across these data sets.  
Further, the metropolitan statistical area definition looks to typically
generate a larger geographical area in the US than Australia; again leading
to issues of comparability.
<br><br>  

### 3 Density Comparisons
#### 3.1 Density by Country
```{r density by country}
densities <- CalcDensities(all_data, ccode = params$ccode, ppn_est = params$ppn_est,
                           area = params$area, density = params$density, group_by = params$ccode)
PlotDensities(densities, density = params$density, ccode = params$ccode)
```
<br><br>  

##### The Median Statistical Area by Density
```{r SA median density}
# Sorting the data by descending data and calculating the cumulative percentage population to find the median
all_by_density <- CalcCumPercent(all_data, cum_pct = params$cum_pct, metric = params$ppn_est, sort_by = params$density, group_by = params$ccode)
mdn_sa <- FindMedian(all_by_density, cum_pct = params$cum_pct)
show_cols <- c(params$ccode, params$sabbr, params$msa_name, params$sa_code, params$ppn_est, params$area, params$density, params$cum_pct)
ShowTable(TableMedian(mdn_sa, show_cols), params$table_css)
```
<br>  

#### Discussion 2022-06-21
So how dense is Australia vs. USA? Well depends on how you calculate it. The
arithmetic method is weighted by areas in the US with very high density e.g. in
New York and Chicago. The median result is interesting as it show the 50th 
percentile or median person in Australia experiences a higher density than
the US.
<br><br>  

#### 3.2 Statistical Areas by Descending Density
```{r SA desc density}
PlotAllDataByDensity(all_by_density, density = params$density, ccode = params$ccode, cum_pct = params$cum_pct, mdn_sa = mdn_sa)
```
<br><br>  

#### 3.3 The Median Metropolitan Area by Population
```{r MSA median popn}
# Group the data into its metropolitan statistical areas
msas <- GroupData(all_data, ppn_est = params$ppn_est, area = params$area, ppn_pct = params$ppn_pct, density = params$density, group_by = c(params$ccode, "MSA_Code", params$msa_name, "Metro_Micro"))

# Sort by population descending
msas_by_popn <- SortMSAsbyPopn(msas, cum_pct = params$cum_pct, ppn_est = params$ppn_est, ppn_pct = params$ppn_pct, msa_name = params$msa_name, group_by = params$ccode)
mdn_msa_popn <- FindMedian(msas_by_popn, cum_pct = params$cum_pct)
show_cols <- c(params$ccode, params$msa_name, params$ppn_est, params$area, params$density, params$cum_pct)
ShowTable(TableMedian(mdn_msa_popn, show_cols), params$table_css)
```
<br><br>  

#### 3.4 Metroplitan Areas by Descending Population
```{r MSA desc popn}
PlotMSAsByPopn(msas_by_popn, ppn_est = params$ppn_est, ccode = params$ccode, cum_pct = params$cum_pct)
```
<br><br>  

#### 3.5 The Median Metropolitan Area by Simple Density
```{r MSA median density}
msas_by_density <- SortMSAsbyDensity(msas, cum_pct = params$cum_pct, density = params$density, ppn_pct = params$ppn_pct, group_by = params$ccode)
mdn_msa_density <- FindMedian(msas_by_density, cum_pct = params$cum_pct)
show_cols <- c(params$ccode, params$msa_name, params$ppn_est, params$area, params$density, params$cum_pct)
ShowTable(TableMedian(mdn_msa_density, show_cols), params$table_css)
```
<br><br>  

#### 3.6 Metropolitan Areas by Descending Simple Density
```{r MSA desc density}
PlotMSAsByDensity(msas_by_density, density = params$density, ccode = params$ccode, cum_pct = params$cum_pct)
```
<br><br>  

## Source Data
### Australia
Data sourced from the Australian Bureau of Statistics. Population data sourced
at the Statistical Area 2 level. See references below for further detail on
Australian statistical geographic standards. Density is given by the ABS but
is recalculated here as the population-area-density numbers do not line up as
given. Data is output to `au.csv` in the data directory.

#### References:
* [Description of the Australian Statistical Geography Standard](https://www.abs.gov.au/statistics/standards/australian-statistical-geography-standard-asgs-edition-3/jul2021-jun2026#asgs-diagram)
* [ABS population data by SA2](https://www.abs.gov.au/statistics/people/population/regional-population/2020-21#data-download)
* [Significant Urban Area to SA2](https://www.abs.gov.au/AUSSTATS/abs@.nsf/DetailsPage/1270.0.55.004July%202016?OpenDocument)

The repo for this code is [here](https://github.com/qwertytam/geo.popn.density)
