---
title: "Population Densities - Comparing Across (Selected) Countries"
output:
  github_document: default
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
keep_md: true
---

## Notebook Summary 
This notebook looks at population density metrics for a set of selected countries.
The basic flow is reading in the data, some data manipulation, then reviewing
various plots

Current countries in the data source:
- Australia
- United States of America

---
  
### 1. Initialize and Read in the Data
Data is sourced from the relevant national statistical agencies and gathered
using scripts in the `data-raw` directory with results stored as `csv` files in
the `inst/extdata` directory and as `rda` files in the `data` directory.

```{r setup}
# Load library for successful knit
library(geo.popn.density)

# Set up variables we will reuse in the notebook
ccode = "Country_Code"
msa_name = "MSA_Name"
ppn_est = "Popn_Est"
area = "Area_km2"
density = "Density"
ppn_pct = "Perc_Popn"
cum_pct = "Cum_Popn_Perc"
method = "method"

# Get the data
all_data <- GetAllData()
```

### 2. Data Review
#### 2.1 Density Calculations by Country
```{r by country}
densities <- CalcDensities(all_data, ccode = ccode, ppn_est = ppn_est,
                               area = area, density = density, group_by = ccode)
PlotDensities(densities, density = density, ccode = ccode)
```

#### 2.2 The Median Statistical Area by Density
```{r SA median density}
# Sorting the data by descending data and calculating the cumulative percentage population to find the median
all_by_density <- CalcCumPercent(all_data, cum_pct = cum_pct, metric = ppn_est, sort_by = density, group_by = ccode)
mdn_sa <- FindMedian(all_by_density, cum_pct = cum_pct)
dplyr::select(mdn_sa, Country_Code, State_Abbr, MSA_Name, SA_Code, Popn_Est, Area_km2, Density, Cum_Popn_Perc)
```

#### 2.3 Statistical Areas by Descending Density
```{r SA desc density}
PlotAllDataByDensity(all_by_density, density = density, ccode = ccode, cum_pct = cum_pct)
```

#### 2.4 The Median Metropolitan Area by Population
```{r MSA median popn}
# Group the data into its metropolitan statistical areas
msas <- GroupData(all_data, ppn_est = ppn_est, area = area, ppn_pct = ppn_pct,
                  density = density,
                  group_by = c(ccode, "MSA_Code", msa_name, "Metro_Micro"))

# Sort by population descending
msas_by_popn <- SortMSAsbyPopn(msas, cum_pct = cum_pct, ppn_est = ppn_est, ppn_pct = ppn_pct, msa_name = msa_name, group_by = ccode)
mdn_msa_popn <- FindMedian(msas_by_popn, cum_pct = cum_pct)
dplyr::select(mdn_msa_popn, Country_Code, MSA_Name, Popn_Est, Area_km2, Density, Cum_Popn_Perc)
```

#### 2.5 Metroplitan Areas by Descending Population
```{r MSA desc popn}
PlotMSAsByPopn(msas_by_popn, ppn_est = ppn_est, ccode = ccode, cum_pct = cum_pct)
```

#### 2.6 The Median Metropolitan Area by Simple Density
```{r MSA median density}
msas_by_density <- SortMSAsbyDensity(msas, cum_pct = cum_pct, density = density, ppn_pct = ppn_pct, group_by = ccode)
mdn_msa_density <- FindMedian(msas_by_density, cum_pct = cum_pct)
dplyr::select(mdn_msa_density, Country_Code, MSA_Name, Popn_Est, Area_km2, Density, Cum_Popn_Perc)
```

#### 2.7 Metropolitan Areas by Descending Simple Density
```{r MSA desc density}
PlotMSAsByDensity(msas_by_density, density = density, ccode = ccode, cum_pct = cum_pct)
```

## Source Data
### Australia
Data sourced from the Australian Bureau of Statistics. Population data sourced
at the Statistical Area 2 level. See references below for further detail on
Australian statistical geographic standards. Density is given by the ABS but
is recalculated here as the population-area-density numbers do not line up as
given. Data is output to `au.csv` in the data directory.

#### References:
* [Description of the Australian Statistical Geography Standard](https://www.abs.gov.au/statistics/standards/australian-statistical-geography-standard-asgs-edition-3/jul2021-jun2026#asgs-diagram)
* [ABS population data by SA2](https://www.abs.gov.au/statistics/people/population/regional-population/2020-21#data-download)
* [Significant Urban Area to SA2](https://www.abs.gov.au/AUSSTATS/abs@.nsf/DetailsPage/1270.0.55.004July%202016?OpenDocument)

The repo for this code is [here](https://github.com/qwertytam/geo.popn.density)

