---
title: "Population Densities - Comparing Across (Selected) Countries"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
keep_md: true
---

## Notebook Summary 
This notebook looks at population density metrics for a set of selected countries.
The basic flow is reading in the data, some data manipulation, then reviewing
various plots.

Current countries in the data source:
- Australia
- United States of America

---
  
### 1. Initialize and Read in the Data
Data is sourced from the relevant national statistical agencies and gathered
using scripts in the `data-raw` directory with results stored as `csv` files in
the `inst/extdata` directory and as `rda` files in the `data` directory.

```{r}
# Load library for successful knit
library(geo.popn.density)

# Set up variables we will reuse in the notebook
ccode = "Country_Code"
msa_name = "MSA_Name"
ppn_est = "Popn_Est"
area = "Area_km2"
density = "Density"
ppn_pct = "Perc_Popn"
cum_pct = "Cum_Popn_Perc"
method = "method"

# Get the data
all_data <- GetAllData()
```

### 2. Data Review
#### 2.1 Density Calculations by Country
```{r}
densities <- CalcDensities(all_data, ccode = ccode, ppn_est = ppn_est,
                               area = area, density = density, group_by = ccode)
PlotDensities(densities, density = density, ccode = ccode)
```

